---
title: ggplot2 Graphics
author: Juan Pacheco
date: '2021-05-08'
slug: ggplot
categories:
  - R
tags:
  - ggplot2
  - R Markdown
  - gapminder
---

<div style="text-align: justify">



We gonna use dplyr, gapminder, RColorBrewer, ggplot2 libraries in this example. 

```{r include=FALSE}
library(dplyr)

library(gapminder)

library(RColorBrewer)

library(ggplot2)

```

Let's load gapminder data frame and see colnames. 
```{r}
data("gapminder")
colnames(gapminder)
```


gapminder is a data frame with 9 variables and more than 10 thousand observations. In variables we have country names, continent, year, life expectancy, population and gross domestic product per capita. 

```{r message=FALSE, warning=FALSE}
gapminder %>% head()
```

## Explore Data 

```{r}
glimpse(gapminder)
```

## Mutate and Filter 

For this example we gonna create a new variable called promedio global and our year of analysis will be 2007. Also, let's take a portion of countries to visualize much better the graphics. 

```{r message=FALSE, warning=FALSE}
datos<- gapminder %>% filter(year=="2007" ) %>% arrange((lifeExp)) %>% mutate(promedio_global= mean(lifeExp))

datos2007<- datos[40:50,]
```

## ggplot2 + Themes 

There are some functions that we can add through our plots . In order to make them more informative: 

**brewer.pal:** makes the color palettes from ColorBrewer available as R palettes. [Source 1](https://cran.r-project.org/web/packages/RColorBrewer/RColorBrewer.pdf)

**Themes:** Themes are a powerful way to customize the non-data components of your plots: i.e. titles, labels, fonts, background, gridlines, and legends. [Source 2](https://www.rdocumentation.org/packages/ggplot2/versions/3.3.3/topics/theme)


**geom_point:** The point geom is used to create scatterplots. [Source 3](https://www.rdocumentation.org/packages/ggplot2/versions/3.3.3/topics/geom_point)

**geom_text:** Adds only text to the plot [Source 4](https://www.rdocumentation.org/packages/ggplot2/versions/3.3.3/topics/geom_label)

**geom_segment:** Draws a straight line between points (x, y) and (xend, yend). [Source 5](https://www.rdocumentation.org/packages/ggplot2/versions/3.3.3/topics/geom_segment)

**geom_vline:** This geom allows you to annotate the plot with vertical lines. [Source 6](https://www.rdocumentation.org/packages/ggplot2/versions/0.9.1/topics/geom_vline)

### Examples 

```{r}
# Add a geom_segment() layer
palette <- brewer.pal(5, "RdYlBu")[-(2:4)]
plt_country_vs_lifeExp<-ggplot(datos2007, aes(x = lifeExp, y = country, color = lifeExp)) +
  geom_point(size = 4) +
  geom_segment(aes(xend = 30, yend = country), size = 2) +
  geom_text(aes(label = lifeExp), color = "white", size = 1.5) +  scale_x_continuous("", expand = c(0,0), limits = c(30,90), position = "top") +
  scale_color_gradientn(colors = palette) + labs(title="Highest and lowest life expectancies, 2007", caption="Source: gapminder")

plt_country_vs_lifeExp
```

```{r}
## More adds 
#Define the theme
plt_country_vs_lifeExp +
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_text(color="black"),
        axis.title = element_blank(),
        legend.position= "none")
```


```{r}
# Add a theme 
step_1_themes <-theme(axis.line.y = element_blank(),
      axis.ticks.y = element_blank(),
      axis.text = element_text(color="black"),
      axis.title = element_blank(),
      legend.position= "none")

global_mean<- mean(datos2007$lifeExp)
x_start <- global_mean + 4
y_start <- 5.5
x_end <- global_mean
y_end <- 7.5

plt_country_vs_lifeExp +
  step_1_themes +
  geom_vline(xintercept= global_mean, color="grey40", linetype=3) + 
  annotate(
    "text",
    x = x_start, y = y_start,
    label = "The\nglobal\naverage",
    vjust = 1, size = 3, color = "grey40")
```



```{r}
#asignamos a annotate a una variable para rebajar el codigo
step_3_annotation <- annotate(
  "text",
  x = x_start, y = y_start,
  label = "The\nglobal\naverage",
  vjust = 1, size = 3, color = "grey40")

#Grafico final
plt_country_vs_lifeExp +  
  step_1_themes +
  geom_vline(xintercept = global_mean, color = "grey40", linetype = 3) +
  step_3_annotation +
  annotate(
    "curve",
    x = x_start, y = y_start,
    xend = x_end, yend = y_end,
    arrow = arrow(length= unit(0.2, "cm"), type = "closed"),
    color = "grey40")

```

    
<div/>
