---
title: Data Manipulation
author: Juan Pacheco
date: '2021-05-13'
slug: data manipulation with dplyr
categories:
  - R
tags:
  - dplyr
  - data manipulation
  - credit
---

## R Package

dplyr, ggplot2, plotly, readxl and knitr.

```{r include=FALSE}
librerias<- c("dplyr","ggplot2","plotly","readxl","knitr")

for (lib in librerias) {
  
      library(lib, character.only = TRUE)  
}

```

Functions that were used:

read_excel() for read excel files

filter() in order to make subsets of our data 

mutate() Create new variables 

as.factor() covert a variable in factor 

glimpse similar to str() applied to a dataframe

group_by() takes an existing tbl and converts it into a grouped tbl where operations are performed "by group"

summarise() o summarize() creates a new data frame that will have 
one or more rows for each combination of grouping variables

n() used from within summarise, mutate and filter 

## Import Data
```{r}
datos<-  read_excel(path = "/Users/Juan Alfredo.DESKTOP-ETJ31JC/Documents/Universidad/Termino 2020 2S/R Samples/Pagina/credit.xlsx", sheet = 2)

```

## Explore Data 

```{r echo=TRUE, message=FALSE, warning=FALSE}
 glimpse(datos)
```

## Descriptive Statistic 

```{r message=FALSE, warning=FALSE}

summary(datos)

```

## Handle missing values NA
```{r}

dt<- datos %>% filter(! is.na(.)) 

```

In this way we can't drop Na's because the rows have values. 

```{r}
dt<- datos %>% filter( `Tipo Producto`!="Sobregiro")  
```

If we filter by "Sobregiro" we drop the NA's in this "case". 

# Group By Class 

```{r echo=TRUE}
cl<- dt %>% group_by(Clase) %>% mutate(Clase = as.factor(Clase)) %>% summarise(count= n())

```

```{r}
kable(cl)
```


# Group by Parent & branch bank
```{r echo=TRUE}
pb<-dt %>% group_by(Sucursal) %>% 
  mutate(`Tipo Crédito` = as.factor(`Tipo Crédito`)) %>% summarise(mean_mora=mean(`Días Mora`))
```

```{r}
kable(pb)
```


# Group by Credit types and Branches bank 
```{r echo=TRUE}
cb<- dt %>% group_by(Sucursal,`Tipo Crédito`) %>% summarise(across(c("Monto Original","Días Mora"), ~ mean(.x, na.rm = TRUE)))
```

```{r}
kable(cb)
```

# Average default by days
```{r message=FALSE, warning=FALSE}
tb1<- dt %>% group_by(`Tipo Crédito`, CALIFICACION) %>% mutate(`Tipo Crédito` = as.factor(`Tipo Crédito`), CALIFICACION = (as.factor(CALIFICACION))) %>% summarise(count = n(),mean_mora=mean(`Días Mora`) )


```
# Table 
```{r}
kable(tb1)
```

# Data visualization

```{r include=FALSE}
p1<-dt %>% group_by(Sucursal) %>%
  summarise(obser= n()) %>%
  mutate(Sucursal = as.factor(Sucursal)) %>%
  ggplot(aes(x = Sucursal, y = obser, fill = Sucursal)) +
  geom_bar(stat = 'identity') +
  theme_minimal() +
  labs(x = "Type", y = "Obsr", title = "Branches") +
  scale_y_continuous()
```

```{r}
ggplotly(p1)
```

```{r include=FALSE}
p2<-dt %>% group_by(Sucursal, `Tipo Crédito`) %>%
    summarise(mean_cred= mean(`Monto Original`)) %>%
  mutate(`Tipo Credito` = as.factor(`Tipo Crédito`)) %>%
  ggplot(aes(x = `Tipo Crédito`, y = mean_cred, fill = `Tipo Crédito`)) +
  geom_bar(stat = 'identity') +
  theme_minimal() +
  labs(x = "Credit types", y = "Initial Amount ", title = "Credit") +
  scale_y_continuous()
```

```{r}
ggplotly(p2)
```









